Author disambiguation of large-scale semi-labeled collections of scientific articles

### 2015-08-22 paper.tex
save

% For algorithms
\usepackage{algorithm}
\usepackage{algorithmicx}
\usepackage[noend]{algpseudocode}
\newcommand{\theHalgorithm}{\arabic{algorithm}}

@@

The author disambiguation solution proposed in this work, along with its
enhancements, are evaluated on data extracted from the INSPIRE portal
\citep{gentil2009information}, a digital library for scientific literature in
high-energy physics. Overall, the portal holds more than 1 million
publications, forming in total a set ${\cal S}$ of more than 10 millions
signatures. Out of these, around 13\% have been \textit{claimed} by their
original authors or marked as such by professional curators. Together, they
constitute a trusted set ${\cal C}^\prime$ of 15388 distinct individuals sharing
36340 unique author names spread within 1201763 signatures on 360066
publications. This data covers several decades in time and dozens of author
nationalities worldwide.

@@

Experiments carried out to study the impact of the proposed algorithmic
components and refinements, as described in Section~\ref{methods}, follow a
standard 3-fold cross-validation protocol, using ${\cal S}^\prime$ as
ground-truth dataset. To replicate the $|{\cal S}^\prime| / |{\cal S}| \approx 13\%$ ratio
of claimed signatures with respect to the total set of signatures, as on the
INSPIRE platform, cross-validation folds are constructed by sampling a training
set ${\cal S}_\text{train}^\prime \subseteq {\cal S}^\prime$ of 13\% of claimed
signatures, and by using the remaining signatures ${\cal S}_\text{test}^\prime = {\cal S}^\prime \setminus {\cal S}_\text{train}^\prime$ for testing. Accordingly, ${\cal
C}_\text{train}^\prime = \{ c^\prime \cap {\cal S}_\text{train}^\prime |
c^\prime \in {\cal C}^\prime \}$ represents the partial known
clusters on the training fold, while ${\cal C}_\text{test}^\prime$ are those
used for testing.

As commonly done in author disambiguation research \citep{levin2012citation},
we evaluate the predicted clusters over training data ${\cal
C}_\text{train}^\prime$, testing data ${\cal C}_\text{test}^\prime$ and the
whole data ${\cal C}^\prime$, using both B3 \glnote{add ref} and pairwise \glnote{add ref} precision, recall
and F-measure, as defined below:
\begin{align}
P_\text{B3}({\cal C}, \widehat{\cal C}, {\cal S}) &= \frac{1}{|{\cal S}|} \sum_{s \in {\cal S}} \frac{|c(s) \cap \widehat{c}(s)|}{|\widehat{c}(s)|} \\
R_\text{B3}({\cal C}, \widehat{\cal C}, {\cal S}) &= \frac{1}{|{\cal S}|} \sum_{s \in {\cal S}} \frac{|c(s) \cap \widehat{c}(s)|}{|c(s)|}\\
F_\text{B3}({\cal C}, \widehat{\cal C}, {\cal S}) &= \frac{2 P_\text{B3}({\cal C}, \widehat{\cal C}, {\cal S}) R_\text{B3}({\cal C}, \widehat{\cal C}, {\cal S})}{P_\text{B3}({\cal C}, \widehat{\cal C}, {\cal S}) + P_\text{B3}({\cal C}, \widehat{\cal C}, {\cal S})}\\
P_\text{pairwise}({\cal C}, \widehat{\cal C}) &= \frac{|p({\cal C}) \cap p(\widehat{\cal C})|}{|p(\widehat{\cal C})|}\\
R_\text{pairwise}({\cal C}, \widehat{\cal C}) &= \frac{|p({\cal C}) \cap p(\widehat{\cal C})|}{|p({\cal C})|}\\
F_\text{pairwise}({\cal C}, \widehat{\cal C}) &= \frac{2 P_\text{pairwise}({\cal C}, \widehat{\cal C}) R_\text{pairwise}({\cal C}, \widehat{\cal C})}{P_\text{pairwise}({\cal C}, \widehat{\cal C}) + R_\text{pairwise}({\cal C}, \widehat{\cal C})}
\end{align}
where $c(s)$ (resp. $\widehat{c}(s)$) is the cluster $c \in {\cal C}$ such that
$s \in c$ (resp. the cluster $\widehat{c} \in \widehat{\cal C}$ such that $s
\in \widehat{c}$), and where $p({\cal C}) = \cup_{c \in {\cal C}} \{ (s_1, s_2)
| s_1, s_2 \in c, s_1 \neq s_2 \}$ is the set of all pairs of signatures from
the same clusters in ${\cal C}$. Intuitively, precision evaluates if signatures
are grouped only with signatures from the same true clusters, while recall
measures the extent to which all signatures from the same true clusters are
effectively grouped together. The F-measure is the harmonic mean between these
two quantities. In the analysis below, we rely primarily on B3 measures for discussing
the results, as pairwise measures tend to favor large clusters (because the number of pairs is quadratic with the cluster size),
hence unfairly giving preference to authors with many publications. By contrast,
B3 measures weight clusters linearly with respect to their size.

@@

\textit{Baseline.}

\glnote{Describe baseline}\\
\glnote{Add an uber table of results}

% The starting point of our analysis is an algorithm that blocks the signatures using the last
% name and first initial strategy, samples the pairs in balanced way, uses gradient boosting
% with all features included for learning the similarity function and clusters using
% semi-supervised hierarchical approach.

>>>>

@@

The author disambiguation solution proposed in this work, along with its
enhancements, are evaluated on data extracted from the INSPIRE portal
\citep{gentil2009information}, a digital library for scientific literature in
high-energy physics. Overall, the portal holds more than 1 million
publications, forming in total a set ${\cal S}$ of more than 10 millions
signatures. Out of these, around 13\% have been \textit{claimed} by their
original authors or marked as such by professional curators. Together, they
constitute a trusted set $({\cal S}^\prime, {\cal C}^\prime)$ of 15388 distinct individuals sharing
36340 unique author names spread within 1201763 signatures on 360066
publications. This data covers several decades in time and dozens of author
nationalities worldwide.

@@

Experiments carried out to study the impact of the proposed algorithmic
components and refinements, as described in Section~\ref{methods}, follow a
standard 3-fold cross-validation protocol, using $({\cal S}^\prime, {\cal C}^\prime)$ as
ground-truth dataset. To replicate the $|{\cal S}^\prime| / |{\cal S}| \approx 13\%$ ratio
of claimed signatures with respect to the total set of signatures, as on the
INSPIRE platform, cross-validation folds are constructed by sampling a training
set ${\cal S}_\text{train}^\prime \subseteq {\cal S}^\prime$ of 13\% of claimed
signatures, and by using the remaining signatures ${\cal S}_\text{test}^\prime = {\cal S}^\prime \setminus {\cal S}_\text{train}^\prime$ for testing. Accordingly, ${\cal
C}_\text{train}^\prime = \{ c^\prime \cap {\cal S}_\text{train}^\prime |
c^\prime \in {\cal C}^\prime \}$ represents the partial known
clusters on the training fold, while ${\cal C}_\text{test}^\prime$ are those
used for testing.

As commonly done in author disambiguation research \citep{levin2012citation},
we evaluate the predicted clusters over training data ${\cal
C}_\text{train}^\prime$, testing data ${\cal C}_\text{test}^\prime$ and the
whole data ${\cal C}^\prime$, using both B3 \glnote{add ref} and pairwise \glnote{add ref} precision, recall
and F-measure, as defined below:
\begin{align}
P_\text{B3}({\cal C}, \widehat{\cal C}, {\cal S}) &= \frac{1}{|{\cal S}|} \sum_{s \in {\cal S}} \frac{|c(s) \cap \widehat{c}(s)|}{|\widehat{c}(s)|} \\
R_\text{B3}({\cal C}, \widehat{\cal C}, {\cal S}) &= \frac{1}{|{\cal S}|} \sum_{s \in {\cal S}} \frac{|c(s) \cap \widehat{c}(s)|}{|c(s)|}\\
F_\text{B3}({\cal C}, \widehat{\cal C}, {\cal S}) &= \frac{2 P_\text{B3}({\cal C}, \widehat{\cal C}, {\cal S}) R_\text{B3}({\cal C}, \widehat{\cal C}, {\cal S})}{P_\text{B3}({\cal C}, \widehat{\cal C}, {\cal S}) + P_\text{B3}({\cal C}, \widehat{\cal C}, {\cal S})}\\
P_\text{pairwise}({\cal C}, \widehat{\cal C}) &= \frac{|p({\cal C}) \cap p(\widehat{\cal C})|}{|p(\widehat{\cal C})|}\\
R_\text{pairwise}({\cal C}, \widehat{\cal C}) &= \frac{|p({\cal C}) \cap p(\widehat{\cal C})|}{|p({\cal C})|}\\
F_\text{pairwise}({\cal C}, \widehat{\cal C}) &= \frac{2 P_\text{pairwise}({\cal C}, \widehat{\cal C}) R_\text{pairwise}({\cal C}, \widehat{\cal C})}{P_\text{pairwise}({\cal C}, \widehat{\cal C}) + R_\text{pairwise}({\cal C}, \widehat{\cal C})}
\end{align}
and where $c(s)$ (resp. $\widehat{c}(s)$) is the cluster $c \in {\cal C}$ such that
$s \in c$ (resp. the cluster $\widehat{c} \in \widehat{\cal C}$ such that $s
\in \widehat{c}$), and where $p({\cal C}) = \cup_{c \in {\cal C}} \{ (s_1, s_2)
| s_1, s_2 \in c, s_1 \neq s_2 \}$ is the set of all pairs of signatures from
the same clusters in ${\cal C}$. Intuitively, precision evaluates if signatures
are grouped only with signatures from the same true clusters, while recall
measures the extent to which all signatures from the same true clusters are
effectively grouped together. The F-measure is the harmonic mean between these
two quantities. In the analysis below, we rely primarily on B3 measures for discussing
the results, as pairwise measures tend to favor large clusters (because the number of pairs is quadratic with the cluster size),
hence unfairly giving preference to authors with many publications. By contrast,
B3 measures weight clusters linearly with respect to their size.

@@

\textit{Baseline.} Results presented in this section are all discussed with
respect to a baseline solution using the following combination of components:
\begin{itemize}
\item Blocking: same surname and the same first given name initial strategy;
\item Linkage function: all 22 features defined in Table~\ref{table:features},
    gradient boosted regression trees as supervised learning algorithm
    and a training set of pairs built by balancing easy and difficult cases.
\item Clustering: agglomerative clustering using average linkage and
    block cuts found to maximize $F_\text{B3}({\cal C}_\text{train}^\prime, \widehat{\cal C}_\text{train}^\prime, {\cal S}_\text{train}^\prime)$.
\end{itemize}

\glnote{Add an uber table of results}
